---
title: "Lightning Talk"
author: "Bryan Stuhr"
output: ioslides_presentation
---
```{R echo=FALSE, message=FALSE, warning=FALSE}
library(NHLData)
library(ggplot2)
library(dplyr)
library(tidyverse)

data(list = data(package = "NHLData")$results[, "Item"])
sch_names <- ls(pattern = "^Sch")
all_nhl <- bind_rows(mget(sch_names))
```

## What is NHLData?

- NHL game scores from every season since 1917
- Over 57,000 games of data
- Contains home/away teams, goals scored, overtime, and playoff info

## Average Goals Per Game Over NHL History
```{r echo=FALSE, message=FALSE, warning=FALSE}
all_nhl$Year <- year(all_nhl$Date)


goals_by_year <- all_nhl %>%
  filter(!is.na(GF), !is.na(GA)) %>%
  group_by(Year) %>%
  summarise(avg_goals = mean(GF + GA, na.rm = TRUE))

ggplot(goals_by_year, aes(x = Year, y = avg_goals)) +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_smooth(se = FALSE, color = "orange", linetype = "dashed") +
  labs(
    title = "Average Goals Per Game Over NHL History",
    x = "Year",
    y = "Average Total Goals",
    caption = "Data: NHLData R package"
  ) +
  theme_minimal()
```

## Home vs Away Scoring Across NHL Eras

```{r echo=FALSE, message=FALSE, warning=FALSE}
all_nhl %>%
  filter(!is.na(GF), !is.na(GA)) %>%
  mutate(Year = year(Date),
         Era = case_when(
           Year < 1940 ~ "Early NHL (1917-1939)",
           Year >= 1940 & Year < 1980 ~ "Mid NHL (1940-1979)",
           Year >= 1980 & Year < 2000 ~ "Modern NHL (1980-1999)",
           Year >= 2000 ~ "Recent NHL (2000-2020)"
         )) %>%
  group_by(Era) %>%
  summarise(Home = mean(GF, na.rm = TRUE),
            Away = mean(GA, na.rm = TRUE)) %>%
  pivot_longer(cols = c(Home, Away), names_to = "Location", values_to = "Avg_Goals") %>%
  ggplot(aes(x = Era, y = Avg_Goals, fill = Location)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("Home" = "darkorange", "Away" = "deepskyblue")) +
  labs(
    title = "Home vs Away Scoring Across NHL Eras",
    x = "",
    y = "Average Goals Per Game",
    caption = "Data: NHLData R package"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 15, hjust = 1))
```

## Most Playoff Wins by Era

```{r echo=FALSE, message=FALSE, warning=FALSE}
all_nhl %>%
  filter(!is.na(Playoff), Playoff == "Y") %>%
  mutate(Year = year(Date),
         Era = case_when(
           Year < 1940 ~ "Early NHL (1917-1939)",
           Year >= 1940 & Year < 1980 ~ "Mid NHL (1940-1979)",
           Year >= 1980 & Year < 2000 ~ "Modern NHL (1980-1999)",
           Year >= 2000 ~ "Recent NHL (2000-2020)"
         ),
         Winner = ifelse(GF > GA, Home, Away)) %>%
  group_by(Era, Winner) %>%
  summarise(Playoff_Wins = n()) %>%
  slice_max(Playoff_Wins, n = 5) %>%
  ggplot(aes(x = reorder(Winner, Playoff_Wins), y = Playoff_Wins, fill = Era)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~Era, scales = "free_y") +
  labs(
    title = "Top 5 Teams with Most Playoff Wins by Era",
    x = "",
    y = "Playoff Wins",
    caption = "Data: NHLData R package"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Conclusions

- The NHL has become more defensive over 100 years
- Home ice advantage has been consistent across all eras
- Montreal Canadiens dominate early/mid NHL playoff wins

## Citations & Next Steps

- Sweet, D.L. (2017). NHLData: Scores for Every Season Since the Founding of the NHL in 1917. 
- R package version 1.0.0. https://CRAN.R-project.org/package=NHLData

- https://github.com/Bryan-Stuhr/Lighting_Talk/tree/master
