---
title: "NHLData Vignette"
author: "Bryan Stuhr"
output: html_document
---

## Introduction

The `NHLData` package contains scores for every NHL game since the league 
was founded in 1917. This vignette walks through how to install the package, 
load the data, and explore it with some basic examples.

The examples in this vignette were written independently using the NHLData package documentation as a refrence. 
The plot code and analysis are orginial work.

```{r eval=FALSE}
install.packages("NHLData")
```

## Loading the Data

```{r message=FALSE, warning=FALSE}
library(NHLData)
library(tidyverse)

data(list = data(package = "NHLData")$results[, "Item"])
sch_names <- ls(pattern = "^Sch")
all_nhl <- bind_rows(mget(sch_names))
```

## What Does the Data Look Like?

```{r}
glimpse(all_nhl)
```

## Example 1 - Average Goals Per Game Over Time

```{r message=FALSE, warning=FALSE}
all_nhl %>%
  mutate(Year = year(Date)) %>%
  filter(!is.na(GF), !is.na(GA)) %>%
  group_by(Year) %>%
  summarise(avg_goals = mean(GF + GA, na.rm = TRUE)) %>%
  ggplot(aes(x = Year, y = avg_goals)) +
  geom_line(color = "steelblue") +
  labs(title = "Average Goals Per Game Over NHL History",
       x = "Year", y = "Average Total Goals")
```

## Example 2 - Home vs Away Scoring

```{r message=FALSE, warning=FALSE}
all_nhl %>%
  filter(!is.na(GF), !is.na(GA)) %>%
  summarise(Home = mean(GF), Away = mean(GA)) %>%
  pivot_longer(everything(), names_to = "Location", values_to = "Avg_Goals") %>%
  ggplot(aes(x = Location, y = Avg_Goals, fill = Location)) +
  geom_col(width = 0.5) +
  labs(title = "Home vs Away Scoring Across NHL History",
       x = "", y = "Average Goals Per Game")
```

## Citation

Sweet, D.L. (2017). NHLData: Scores for Every Season Since the Founding 
of the NHL in 1917. R package version 1.0.0. 
https://CRAN.R-project.org/package=NHLData
